CUBE: A Memory-lite Cosmological $N$-body Algorithm

### 2017-10-06 ms_caf/ms_caf.tex
initial conditions rewritten

The first step is the calculation of the potential field. At a sufficient high initial redshift $z_i$, we firstly generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space, where $T(k)$ is given by the assumed cosmological model. Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by a Poisson equation $-\nabla^2\Phi=\delta_L$. At this step, a global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field into the final displacement potential $\Phi$. The initial condition generator is applied only once, and in principle, at some costs of speed, one can use only one scalar field on the fine mesh, and FFTs can be done in-place. If the number of particles equals the number of fine grids, then the memory consumption is dominated by the single precision fine mesh field, or 4 bpp.

The second step is to generate particles on Lagrangian coordinates and displace them with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}. The particles are displaced by locally differentiating $\Phi$ in real space meanwhile their initial velocities are assigned and immediately converted into integer-based velocity by Equation (\ref{eq.nu}). PID.

The particles can be generated and saved to disk tile by tile, locally. Assuming that we initially put particles at the center of every fine grid, we first generate these regularly placed particles in an extended tile, and then according to their new positions (coarse grid positions) given by ZA we get a coarse density field on tile. Secondly, according to the existing coarse density field, we apply ZA again to get their accurate positions and initial velocities, and we generate their integer based phase space coordinates and put them in correct orders according to the coarse density field. Lastly, we discard those particles out of the physical region of the tile, and write only particles in physical region on to the disk. As we will see in Section \ref{sss.algorithm}, {\tt update\_xp} paragraph, it is very similar to the ``drift'' step in the main $N$-body code, where an ``buffered'' state of particles is firstly created by two iterations of particles, and the buffered particles are discarded to achieve a ``disjoint'' state. After working on this tile, the arrays for positions and velocities are freed up and can be reused for the next tile. During this step, the only major memory usage is given by the fine grid potential $\Phi$.

>>>>

The first step is the calculation of the potential field. At a sufficient high initial redshift $z_i$, we firstly generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space, where $T(k)$ is given by the assumed cosmological model. Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by a Poisson equation $-\nabla^2\Phi=\delta_L$. At this step, global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field into the final displacement potential $\Phi$. The initial condition generator is applied only once, and in principle, at some costs of speed, one can use only one scalar field on the fine mesh, and FFTs can be done in-place. If the number of particles equals the number of fine grids, then the memory consumption is dominated by the single precision fine mesh field, or 4 bpp.

The second step is to generate particles on Lagrangian coordinates ${\bs q}$ and displace them with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}, by locally differentiating $\Phi$ in real space. There are two iterations over particles' Lagrangian coordinates ${\bs q}$ (usually at the center of every fine grid). The first iteration calculates particles' new, coarse grid positions by ZA and get a coarse density field on tile. The second iteration, ZA is applied again to calculate the accurate positions of particles and their velocities, and convert them into integer based ones by Equations (\ref{eq.chi},\ref{eq.nu}) and place them in a certain order according to the coarse density field generated by the first iteration. Lastly, we discard those particles out of the physical region of the tile, and write only particles in physical region on to the disk. As we will see in Section \ref{sss.algorithm}, {\tt update\_xp} paragraph, it is very similar to the ``drift'' step in the main $N$-body code, where an ``buffered'' state of particles is firstly created by two iterations of particles, and the buffered particles are discarded to achieve a ``disjoint'' state. If PIDs are needed, they are generated at the second iteration and are manipulated similarly as integer based velocities. Because particles are generated and saved to disk tile by tile, locally, after working on certain tile, the arrays for positions, velocities and PIDs (if any) are freed up and can be reused for the next tile. During this step, the only major memory usage is still dominated by the fine grid potential $\Phi$.

After working on all tiles, we sum up the total particle number on this image, {\tt nplocal}, and write it in the header of the checkpoint. Finally, $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm optional})\}$ is rewritten to the disk as the checkpoint of $z=z_i$.
